mathml: Translate R expressions to MathML and LaTeX/MathJax

### 2022-11-30 Draft1.Rmd
update 30/11

---
title: "mathml: Translate R expressions to MathML and LaTeX/MathJax"
date: "2022-11-09"
abstract: >
  This R package translates R objects to suitable entities in MathML or LaTeX,
  thereby enabling pretty mathematical representation of R functions or
  algorithms in data analysis, scientific reports and interactive web content.
  R code text descriptions and mathematical content already coexist,
  side-by-side, in the RMarkdown document rendering language. The present
  package allows to use the same R\ objects for both data analysis and
  typesetting in documents or web content. By tightening the link between the
  statistical analysis and its verbal or symbolic description, this new feature
  is a futher step towards reprocible science. The package can be extended by
  user-defined hooks that allow to replace variable names and new R functions by
  a suitable symbolic representation. We provide two working examples that
  illustrate the use of the new package in a scientific article and in an
  interactive web page.
draft: true
author:  
  # see ?rjournal_article for more information
  - name: Matthias Gondan
    affiliation: University of Innsbruck
    address:
    - Department of Psychology
    - Innsbruck, Austria
    url: https://www.britannica.com/animal/quokka
    orcid: 0000-0001-9974-0057
    email:  Matthias.Gondan-Rochon@uibk.ac.at
  - name: Irene Alfarone
    url: https://www.britannica.com/animal/bilby
    email: Irene.Alfarone@uibk.ac.at
    orcid: 0000-0002-8409-8900
    affiliation: University of Innsbruck
    address:
    - Department of Psychology
    - Innsbruck, Austria
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(plotly)
library(ggplot2)
library(palmerpenguins)
```

@@

The paper is organized as follow. Section 2 presents the background of the mathml package, namely Prolog and the R package `rolog`. Section 3 illustrates the features of the mathml package, providing some examples. Section 4 embeds the Discussion and the presentation of a Case Study. Section 5 is devoted to Conclusions and Further Research. The reader may also find some additional characteristics of the package and its implementation in the Appendix.

@@

The automatic translation of R expressions to mathematical output is achieved through a Prolog interpreter. Prolog is a classical logic programming language with many applications in expert systems, computer linguistics and traditional, that is, symbolic artificial intelligence. The main strength of Prolog is its concise representation of facts and rules for the representation of knowledge and grammar, as well as its efficient built-in search engine for closed world domains. R, as it is well-known, is a statistical programming language for data analysis and statistical modeling which is widely used in academia and industry. Besides the core library, a lot of packages have been developed for all kinds of statistical problems, including statistics-based artificial intelligence tools such as neural networks for machine learning and deep learning. Whereas Prolog is weak in statistical computation, but strong in symbolic manipulation, the converse may be said for the R language. The `rolog` package then bridges this gap by providing an interface to a SWI-Prolog distribution in R. The communication between the two systems is mainly in the form of queries from R to Prolog, but two predicates allow Prolog to ring back and evaluate terms in R. 
Here is reported a first trivial example. At the Prolog end, there is a handler for `pbinom/3` that translates the term into a pretty MathML syntax like P_bi(X <= k; N, pi).

```{r, echo=TRUE}
library(rolog)
library(mathml)
term = quote(pbinom(k, N, p))

@@

_I encounter a problem, when I explicit the 'asis' in the results='', the execution of the Rmd is halted. Yet, only when I use mathml()._

@@

```{r, echo = TRUE}
term = quote(integrate(sin, 0L, 2L*pi))
mathml(term)
eval(term)
```

@@

    # Recurse into arguments
    term[-1] <- lapply(term[-1], canonical)
  }
  return(term)
}
# A custom function
g <- function(u)
{
  sin(u)
}
# # Mixture of (partially) named and positional arguments in unusual order
term <- quote(2L * integrate(low=-Inf, up=Inf, g)$value)
# mathml(canonical(term))
# # It is a bit of a mystery that R knows the result of this integral.
# eval(term)
```

@@

# Customizing tooltip design with \pkg{ToOoOlTiPs}

\pkg{ToOoOlTiPs} is a packages for customizing tooltips in interactive graphics, it features these possibilities.

# A gallery of tooltips examples

The \CRANpkg{palmerpenguins} data [@palmerpenguins]  features three penguin species which has a lovely illustration by Alison Horst in Figure \@ref(fig:penguins-alison).

```{r penguins-alison, out.width = "100%", out.height = "30%", fig.cap = "Artwork by \\@allison\\_horst", fig.alt="A picture of three different penguins with their species: Chinstrap, Gentoo, and Adelie. "}
# knitr::include_graphics("penguins.png")
```

Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:penguins-tab-interactive)', '\\@ref(tab:penguins-tab-static)'))` prints at the first few rows of the `penguins` data:

```{r penguins-tab-interactive, eval = knitr::is_html_output(), layout = "l-body-outset"}
knitr::kable(head(penguins), format = "html", caption = "A basic table")
```

```{r penguins-tab-static, eval = knitr::is_latex_output()}
knitr::kable(head(penguins), format = "latex", caption = "A basic table") %>% 
  kableExtra::kable_styling(font_size = 7)
```

Figure `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(fig:penguins-plotly)', '\\@ref(fig:penguins-ggplot)'))` shows an `r ifelse(knitr::is_html_output(), 'interactive', '')` plot of the penguins data, made using the `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\CRANpkg{plotly}', '\\CRANpkg{ggplot2}'))` package.

```{r penguins-plotly, echo = TRUE, fig.height = 5, fig.cap="A basic interactive plot made with the plotly package on palmer penguin data. Three species of penguins are plotted with bill depth on the x-axis and bill length on the y-axis. When hovering on a point, a tooltip will show the exact value of the bill depth and length for that point, along with the species name.", include=knitr::is_html_output(), eval=knitr::is_html_output(), fig.alt = "A scatterplot of bill length against bill depth, both measured in millimetre. The three species are shown in different colours and loosely forms three clusters. Adelie has small bill length and large bill depth, Gentoo has small bill depth but large bill length, and Chinstrap has relatively large bill depth and bill length."}
p <- penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, 
             color = species)) + 
  geom_point()
ggplotly(p)
```

```{r penguins-ggplot, echo = TRUE, fig.height = 5, fig.cap="A basic non-interactive plot made with the ggplot2 package on palmer penguin data. Three species of penguins are plotted with bill depth on the x-axis and bill length on the y-axis. Visit the online article to access the interactive version made with the plotly package.", include=knitr::is_latex_output(), eval=knitr::is_latex_output()}
penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, 
             color = species)) + 
  geom_point()
```

# Summary

We have displayed various tooltips that are available in the package \pkg{ToOoOlTiPs}.

>>>>

---
title: "mathml: Translate R expressions to MathML and LaTeX/MathJax"
date: "2022-11-09"
abstract: >
  This R package translates R objects to suitable entities in MathML or LaTeX,
  thereby enabling pretty mathematical representation of R functions or
  algorithms in data analysis, scientific reports and interactive web content.
  R code text descriptions and mathematical content already coexist,
  side-by-side, in the RMarkdown document rendering language. The present
  package allows to use the same R\ objects for both data analysis and
  typesetting in documents or web content. By tightening the link between the
  statistical analysis and its verbal or symbolic description, this new feature
  is a futher step towards reproducible science. The package can be extended by
  user-defined hooks that allow to replace variable names and new R functions by
  a suitable symbolic representation. We provide two working examples that
  illustrate the use of the new package in a scientific article and in an
  interactive web page.
draft: true
author:  
  # see ?rjournal_article for more information
  - name: Matthias Gondan
    affiliation: University of Innsbruck
    address:
    - Department of Psychology
    - Innsbruck, Austria
    url: https://www.britannica.com/animal/quokka
    orcid: 0000-0001-9974-0057
    email:  Matthias.Gondan-Rochon@uibk.ac.at
  - name: Irene Alfarone
    url: https://www.britannica.com/animal/bilby
    email: Irene.Alfarone@uibk.ac.at
    orcid: 0000-0002-8409-8900
    affiliation: University of Innsbruck
    address:
    - Department of Psychology
    - Innsbruck, Austria
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(plotly)
library(ggplot2)
library(palmerpenguins)

catmath <- function(term)
{
  s <- substitute(term)
  c <- canonical(s)
  mathml(c)
}

catmathjax <- function(term)
{
  s <- substitute(term)
  c <- canonical(s)
  mathjax(c)
}

```

@@

The paper is organized as follow. Section 2 presents the background of the mathml package, namely Prolog and the R package \pkg{rolog}. Section 3 illustrates the features of the \pkg{mathml} package, providing some examples. Section 4 embeds the Discussion and the presentation of a Case Study. Section 5 is devoted to Conclusions and Further Research. The reader may also find some additional characteristics of the package and its implementation in the Appendix.

@@

The automatic translation of R expressions to mathematical output is achieved through a Prolog interpreter. Prolog is a classical logic programming language with many applications in expert systems, computer linguistics and traditional, that is, symbolic artificial intelligence. The main strength of Prolog is its concise representation of facts and rules for the representation of knowledge and grammar, as well as its efficient built-in search engine for closed world domains. R, as it is well-known, is a statistical programming language for data analysis and statistical modeling which is widely used in academia and industry. Besides the core library, a lot of packages have been developed for all kinds of statistical problems, including statistics-based artificial intelligence tools such as neural networks for machine learning and deep learning. Whereas Prolog is weak in statistical computation, but strong in symbolic manipulation, the converse may be said for the R language. The \pkg{rolog} package then bridges this gap by providing an interface to a SWI-Prolog distribution in R. The communication between the two systems is mainly in the form of queries from R to Prolog, but two predicates allow Prolog to ring back and evaluate terms in R. 
Here is reported a first trivial example. At the Prolog end, there is a handler for `pbinom/3` that translates the term into a pretty MathML syntax like P_bi(X <= k; N, pi).

```{r, echo=TRUE, results='asis', eval=knitr::is_html_output()}
library(rolog)
library(mathml)
term = quote(pbinom(k, N, p))

@@

```{r, echo=TRUE, results='asis', eval=knitr::is_latex_output()}
library(rolog)
library(mathml)
term = quote(pbinom(k, N, p))

# Pretty print
mathjax(term)
```

@@

```{r,  results = 'asis', echo = TRUE, eval=knitr::is_html_output()}
term = quote(integrate(sin, 0L, 2L*pi))
mathml(term)
```

```{r, echo = TRUE}
eval(term)
```

@@

    # Recurse into arguments
    term[-1] <- lapply(term[-1], canonical)
  }
  return(term)
}
# A custom function
g <- function(u)
{
  sin(u)
}
# # Mixture of (partially) named and positional arguments in unusual order
term <- quote(2L * integrate(low=-Inf, up=Inf, g)$value)
# mathml(canonical(term))
# # It is a bit of a mystery that R knows the result of this integral.
eval(term)
```

@@

# Package capabilities

\pkg{mathml} is a package for pretty mathematical representation of R functions or algorithms in data analysis, scientific reports and interactive web content. The features currently supported are listed below, following the order proposed by [@murrell2000].

## Arithmetic operators
The \pkg{mathml} package provides the graphical representation for addition, subtraction, multiplication and division. Additionally, we have provided two possible ways to represent multiplication: With sign and by juxtaposition and two different kinds of division.

```{r, results="asis", echo=TRUE,  eval=knitr::is_html_output()}
catmath(x + y)
catmathjax(x + y)

catmath(x - y)
catmathjax(x - y)

# Plus Minus
catmath(x %+-% y)
catmathjax(x %+-% y)

catmath(x*y)
catmathjax(x*y)

# %*%
catmath(x %*% y)
catmathjax(x %*% y)

# %.%
catmath(x %.% y)
catmathjax(x %.% y)

catmath(X %x% Y)
catmathjax(X %x% Y)

catmath(x/y)
catmathjax(x/y)

# %/%: integer division
catmath(x %/% y)
catmathjax(x %/% y)
```

## Fractions
As in [@murrell2000] the `frac` and `over` functions aim to represent fractions. `choose` and `lchoose`, as in R\ \pkg{base}, display the binomial coefficient.

```{r, results="asis", echo=TRUE,  eval=knitr::is_html_output()}
catmath(frac(x, y))
catmathjax(frac(x, y))

catmath(over(x, y))
catmathjax(over(x, y))

catmath(choose(n, k))
catmathjax(choose(n, k))

catmath(choose(k=K, n=N))
catmathjax(choose(k=K, n=N))

catmath(lchoose(n, k))
catmathjax(lchoose(n, k))
```

## Relations

Among the possible relational operators, following [@murrell2000], we included:

```{r, results="asis", echo=TRUE,  eval=knitr::is_html_output()}
catmath(x==y)
catmathjax(x==y)

catmath(x < y)
catmathjax(x < y)

catmath(x <= y)
catmathjax(x <= y)

catmath(x >= y)
catmathjax(x >= y)

catmath(!x)
catmathjax(!x)

catmath(x != y)
catmathjax(x != y)

catmath(X ~ Y)
catmathjax(X ~ Y)

catmath(x %~~% y)
catmathjax(x %~~% y)

catmath(x %==% y)
catmathjax(x %==% y)

catmath(x %=~% y)
catmathjax(x %=~% y)

catmath(x %prop% y)
catmathjax(x %prop% y)

```

## Set relations

Here is provided a series of useful set relations.

_Still to be implemented_

```{r, results="asis", echo=TRUE,  eval=knitr::is_html_output()}
catmath(a %in% A)
catmathjax(a %in% A)

catmath(intersect(A, B))
catmathjax(intersect(A, B))

catmath(union(A, B))
catmathjax(union(A, B))

catmath(is.null(intersect(A, B)))
catmathjax(is.null(intersect(A, B)))
```

## Arrows

Following once again [@murrell2000], we provide here a list of useful arrows

```{r, results="asis", echo=TRUE,  eval=knitr::is_html_output()}
# x %<->% y
catmath(x %<->% y)
catmathjax(x %<->% y)

# x %->% y
catmath(x %->% y)
catmathjax(x %->% y)

# x %<-% y
catmath(x %<-% y)
catmathjax(x %<-% y)

# x %up% y:
catmath(x %up% y)
catmathjax(x %up% y)

# x %down% y:
catmath(x %down% y)
catmathjax(x %down% y)

# x %<=>% y: solved it using iff
catmath(x %<=>% y)
catmathjax(x %<=>% y)

# x %=>% y: solved using rArr
catmath(x %=>% y)
catmathjax(x %=>% y)

# x %<=% y: solved using lArr
catmath(x %<=% y)
catmathjax(x %<=% y)

# x %dblup% y:
catmath(x %dblup% y)
catmathjax(x %dblup% y)

# x %dbldown% y:
catmath(x %dbldown% y)
catmathjax(x %dbldown% y)

```

# A case study

For a more clear representation of what the \pkg{mathml} package is supposed to do, we provide here a case study, based on the work by Schwarz [-@schwarz1994], in which the features of \pkg{mathml} are implemented. The reader
may recognize that Schwarz [-@schwarz1994]'s study can be traced back to the strand of
decision-making models, where decision-making is assumed to be a process of
noisy accumulation of information over time [e.g., @ratcliff2016]. The aim of
Schwarz' [-@schwarz1994] study is to present a new explanation of redundancy gains when
observers make speeded responses to stimuli of different sources, and the same
information is presented on two channels. Schwarz [-@schwarz1994] describes a model that assumes
the superposition of channel-specific diffusion processes that eventually reach
an absorbing barrier to elicit the response. For a detailed description the
reader may refer to the original article.

```{r, results="asis", echo=FALSE,  eval=knitr::is_html_output()}

# Improve display
hook(quote(mu_A), quote(mu["A"]))
hook(quote(mu_B), quote(mu["B"]))
hook(quote(sigma_A), quote(sigma["A"]))
hook(quote(sigma_B), quote(sigma["B"]))
hook(quote(mu_M), quote(mu["M"]))
hook(quote(zn), quote(z[tau]))
hook(quote(M), quote(overline(X)))
```

### The model

Schwarz' [-@schwarz1994] model refers to two stimuli A and B, presented either
alone or in combination (redundant stimuli), with the redundant stimuli being
presented either simultaneously or with onset asynchrony \(\tau\). The channel
activation is described as a two-dimensional Wiener process with
drifts \(\mu_\mathrm A\), \(\mu_\mathrm B\),
variances \(\sigma^2_\mathrm A\), \(\sigma^2_\mathrm B\),
correlation \(\rho_\mathrm{AB}\) and initial
conditions \(X_i(t=0)=0, i=\mathrm{A, B}\). The response is elicited when the
process reaches the absorbing barrier \(c > 0\) for the first time. In combined
stimuli ("redundant" stimuli), the overall diffusion process
is \(X_\mathrm{AB}(t) = X_\mathrm A(t) + X_\mathrm B(t)\) , which is again a
Wiener process with drift \(\mu_\mathrm A + \mu_\mathrm B\) and
variance \(\sigma^2_\mathrm A + \sigma^2_\mathrm B
  + 2\rho_\mathrm{AB}\sigma_\mathrm A\sigma_\mathrm B\).

In single-target trials, the first passages at \(c\) are expected at
\(E\left[D_i\right] = \frac{c}{\mu_i}, i = \mathrm{A, B}\), whereas in synchronous
redundant-target trials, we have
\(E[D_\mathrm{AB}]= \frac{c}{\mu_\mathrm A + \mu_\mathrm B}\).

For asynchronous stimuli, Schwarz [-@schwarz1994] derived the expected
first-passage time \(E[D_{\mathrm A(\tau)\mathrm B}]\) as a function of the
stimulus onset asyncrony \(\tau > 0\) (Eq.\ 10),

```{r, results="asis", echo=TRUE,  eval=knitr::is_html_output()}
f <- function(tau, c, mu_A, sigma_A, mu_B, sigma_B)
{ dfrac(c, mu_A) + (dfrac(1L, mu_A) - dfrac(1L, mu_A + mu_B)) *
    ((mu_A*tau - c) * pnorm(dfrac(c - mu_A*tau, sqrt(sigma_A^2L*tau)))
      - (mu_A*tau + c) * exp(dfrac(2L*mu_A*c, sigma_A^2L))
        * pnorm(dfrac(-c - mu_A*tau, sqrt(sigma_A^2L*tau))))
}

mathml(f)
```

The observable response time is assumed to be the sum of \(D\), the time
employed to reach the threshold for the decision, and \(M\), denoting the of the
other processes.

\(E[T]= E[D + M] = E[D] + E[M] = E[D] + \mu_M\)

### Methods

Schwarz [-@schwarz1994] applied the model to Miller’s [-@miller1986] data from
a redundant signals task with 13 onset
asynchronies \(0, \pm33, \pm67, \pm100, \pm133, \pm167, \pm\infty\) ms,
where 0 refers to synchronous AB, \(\infty\) refers to the single-target
presentation, and negative \(\tau\) denote those conditions in which B is
presented before A. Each condition was replicated 400 times. The 13 mean RTs and
standard deviations are given in Table 2 below.

Then, from Schwarz' [-@schwarz1994] Equations [2] [3] and [10], we calculated the expected first time passage for the respective \(\tau\) and the parameters \(\mu_ \mathrm A = 0.53, \mu_ \mathrm B = 1.32, \sigma_ \mathrm A = 4.43, \sigma_ \mathrm B = 18.3, c = 100, \mu_\mathrm M = 161\)

````{r, results="asis", echo=FALSE,  eval=knitr::is_html_output()}
mu_A = 0.53
mu_B = 1.32
sigma_A = 4.4
sigma_B = 18.3
tau = c(-Inf, -167, -133, -100, -67, -33, 0, 33, 67, 100, 133, 167, Inf)
c = 100
mu_M = 161
````

```{r, results="asis", echo=FALSE,  eval=knitr::is_html_output()}

# Conditions

g <- function(tau, c, mu_A, sigma_A, mu_B, sigma_B)
{
    if( tau == Inf) return(dfrac(c, mu_A));
    if( tau == -Inf) return(dfrac(c, mu_B));
    if( tau > 0L)  return(f(tau, c, mu_A, sigma_A, mu_B, sigma_B));
    if( tau == 0L) return(dfrac(c, mu_A + mu_B));
    if( tau < 0L) return(f(-tau, c,  mu_B, sigma_B, mu_A, sigma_A))
}

mathml(g)

```

For instance, we can see that with \(\tau = -167\) the expected first passage time is:

```{r, results="asis", echo=FALSE,  eval=knitr::is_html_output()}
mu_A = 0.53
mu_B = 1.32
sigma_A = 4.4
sigma_B = 18.3
tau = -167
c = 100
mu_M = 161

g(tau, c, mu_A, sigma_A, mu_B, sigma_B)

```

To which we have to add \(\mu_M = 161\) (which represents the amount of time that the movement of providing the answer takes) and we obtain:

```{r, results="asis", echo=FALSE,  eval=knitr::is_html_output()}
g(tau, c, mu_A, sigma_A, mu_B, sigma_B) + mu_M
```

Similarly, we obtain the first-time passage for every \(\tau\):

```{r, results="asis", echo=FALSE,  eval=knitr::is_html_output()}

G <- Vectorize(g, vectorize.args = 'tau')

h <- function(tau, c, mu_A, sigma_A, mu_B, sigma_B, mu_M)
{
  G(tau, c, mu_A, sigma_A, mu_B, sigma_B) + mu_M
}
mathml(h)
```

```{r, results="asis", echo=FALSE,  eval=knitr::is_html_output()}

mu_M = 161
mu_A = 0.53
mu_B = 1.32
sigma_A = 4.4
sigma_B = 18.3
tau = c(-Inf, -167, -133, -100, -67, -33, 0, 33, 67, 100, 133, 167, Inf)
c = 100

#h(tau, c, mu_A, sigma_A, mu_B, sigma_B, mu_M)

```

### Model fitting and Parameters Estimation

Then we proceeded with the model fitting, using the data provided by Miller [-@miller1986] study in Schwarz [-@schwarz1994]. First, we calculated the predicted means RT as \(E[RT]=E[T]+\mu_M\).

We briefly remind that in diffusion models there is across-trial variability, and the drift rates are assumed to be normally distributed [@ratcliff2008] with at the first passage of \(c\) \(X(t) \sim N(\mu t , \sigma^2 t)\). Where \(T\) is expressed as \(T= \arg \min_t X(t)\geq c\) with \(T \sim IG (c, \mu, \sigma^2)\) with mean \(E(T)=\displaystyle \frac{c}{\mu}\) and variance \(Var(T)=\displaystyle \frac{c\sigma^2}{\mu^3}\). Furthermore, for the Central Limit Theorem \(\bar{T} = \displaystyle \frac{\Sigma_iT_i}{N} \dot\sim N (\displaystyle \frac{c}{\mu} , \displaystyle \frac{c\sigma^2}{\mu^3})\)

For these reasons, in order to obtain the Goodness of Fit measure we proceeded with a z-standardisation.

```{r, results="asis", echo=FALSE,  eval=knitr::is_html_output()}

zf <-  function(M, s, N, tau, c, mu_A, sigma_A, mu_B, sigma_B, mu_M)
{
  E <- h(tau, c, mu_A, sigma_A, mu_B, sigma_B, mu_M)
  dfrac(M-E, s/sqrt(N))
}

mathml(zf)
```

```{r, results="asis", echo=FALSE,  eval=knitr::is_html_output()}

# The vector X contains the observed values, the vector s their observed standard deviation and N = 400 are the number of trials.

M = c(231, 234, 230, 227, 228, 221, 217, 238, 263, 277, 298, 316, 348)
s = c(56, 58, 40, 40, 32, 28, 28, 28, 26, 30, 32, 34, 92)
N = c(400, 400, 400, 400, 400, 400, 400, 400, 400, 400, 400, 400, 400)

zn <- zf(M, s, N, tau, c, mu_A, sigma_A, mu_B, sigma_B, mu_M)
# zn

```

Then, we calculated the Goodness of Fit as follows:

\(X^2=\Sigma_\tau z^2_\tau \sim\chi^2(13-5\  dgf)\).

```{r, results="asis", echo=FALSE,  eval=knitr::is_html_output()}

par <- c(mu_A = 0.53, sigma_A = 4.4, mu_B = 1.32, sigma_B = 18.3, mu_M = 161)

# Goodness of Fit

gof <- function(par, tau, M, s, N)
{
  zn <- zf(M, s, N, tau, c=100L, mu_A=par["mu_A"], sigma_A=par["sigma_A"], mu_B=par["mu_B"], sigma_B=par["sigma_B"], mu_M=par["mu_M"])
  sum(zn^2L)
}

mathml(gof)

n <- gof(par, tau, M, s, N)
#n

#pchisq(n, 8, lower.tail = FALSE, log.p = FALSE)

```

```{r, results="asis", echo=FALSE,  eval=knitr::is_html_output()}

lsq = function(tau, M, s, N)
{
  start = c(mu_A = 0.3, sigma_A = 2, mu_B = 1 , sigma_B = 5, mu_M = 200)
  optim(start, fn = gof,
    tau = tau, M = M, s = s, N = N, hessian = TRUE, control = list(fnscale = 1))
}

# mathml(lsq)

# Parameters' estimates
fit <-  lsq(tau, M, s, N)

est = fit$par
se = sqrt(diag(solve(fit$hessian)))
cilo = qnorm(0.025, est, se)
ciup = qnorm(0.975, est, se)

# Lower bound
#cilo

# Upper bound
#ciup

```

### Results

The calculated Goodness of Fit value is \(X^2= 30.54\), \(p= 0.0002\). Additionally, we proceeded with the parameters' estimation with the Least Squares Method.

The values of the estimated parameters reported in the table below are consistent with those reported by Schwarz [-@schwarz1994].

: Table 1: Estimated Model's Parameters

| Parameters | Estimate | 95% Confidence Interval
| :---: | :---: | :---:|
| \( \mu_ \mathrm A \)| 0.53 | 0.51, 0.56
| \( \sigma_ \mathrm A \)| 4.4 | 3.8, 5.1
| \( \mu_ \mathrm B \)| 1.32 | 1.2, 1.4
| \( \sigma_ \mathrm B \)| 18.3 | 11.6, 24.9
| \( \mu_ \mathrm M \)| 161 | 157, 165

Furthermore, we can observe from the values reported in the Table below that, with few exceptions (e.i. \(\tau= 100\)) Miller's [-@miller1986] observed values are quite similar to the one predicted by Schwarz' [-@schwarz1994] model. Showing that, despite the calculated Goodness of Fit measure, the model proposed by Schwarz [-@schwarz1994] sufficiently fit the data.

: Table 2: Predicted Means RT and Observed Means RT and SDs from Miller's [-@miller1986] study in Schwarz [-@schwarz1994]

| SOA      | Predicted Mean RT | Observed Mean RT | Observed SD |
|:--------:|:-----------------:|:----------------:|:-----------:|
\(-\infty\)| 237 | 231 | 56 |
\(-167\)   | 231 | 234 | 58 |
\(-133\)   | 230 | 230 | 40 |
\(-100\)   | 228 | 227 | 40 |
\(-67\)    | 226 | 228 | 32 |
\(-33\)    | 222 | 221 | 28 |
\(0\)      | 215 | 217 | 28 |
\(33\)     | 239 | 238 | 28 |
\(67\)     | 262 | 263 | 26 |
\(100\)    | 283 | 277 | 30 |
\(133\)    | 300 | 298 | 32 |
\(167\)    | 313 | 316 | 34 |
\(\infty\) | 350 | 348 | 92 |

 _Note_.  Parameters: \(\mu_ \mathrm A = 0.53, \mu_ \mathrm B = 1.32, \sigma_ \mathrm A = 4.4, \sigma_ \mathrm B = 18.3, c = 100, \mu_\mathrm M = 161\)

# Conclusions

* General introduction on R and its features
* What is mathml 
* issues the user must be aware of
* points of strength
* future directions
* (inspirational ending)
* ways to install rolog and mathml

# Acknowledgments
